Technical Report

### 2017-09-28 fully-abstract-comp/technical-report.tex
Added responses to Dominique's comments

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
\lau{13-09-2017: For now I have added linearity to the seals.}
\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

The source machine is also a capability machine with memory capabilities and sealed capabilities. Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    n & \in & \nats \\
    \src{r} &\in &  \RegName \setminus \{\rstk\}\\
    \src{\rn} &\defbnf & \src{r} \mid n \\
    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
  \end{array}
\]
\lau{30-08-2017: We had $\rret$ as not being one of the registers on the source machine. I could not recall why we had this, so I removed it. (I think it may have been from when we had call and return instructions?)}
\dominique{26-9-2017: I think we should allow $\rstk$ in the source language too.}
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.

@@

\dominique{I think we need to do something about tail calls in the source language?}

@@

\section{Compiler}
\[
\comp{\cdot} : \src{i}^* \fun \trg{i}^*
\]
\dominique{Regarding the type signature for $[\cdot]$ above: does this mean that it maps lists of instructions to lists of instructions? Is there another function that compiles a single instruction to a list of instructions?}

@@

\dominique{Is this example still relevant? Isn't it solved differently by the variable-length-instructions idea?}
This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{}{}$ and $\sreturn$ instructions. The below example is a motivation for having variable length instructions in the source language because if we have enough memory to ``do the same'' in the target language as in the source language, then the below example does not work.

@@

\section{Notes}
\subsection{Leuven stay conclusions}
\begin{enumerate}
\item Enter capabilities replaces by sealed code/data pairs \label{item:first-point}
  \begin{itemize}
  \item To allow us to forbid dynamic code generation
  \end{itemize}
\item Conditional full-abstraction
  \begin{itemize}
  \item No undefined behaviour by \emph{trusted code} in any context implies full-abstraction. (Blame idea: undefined as undef of current pc, so current pc can be checked for whether it was the trusted code).
  \item Avoids dynamic checks to protect trusted code against itself.
  \item Possible due to point \ref{item:first-point}.
  \end{itemize}
\item replace push/pull/call/ret/sload by symbolic return pointer (pair) and stack pointer.
  \begin{itemize}
  \item Allows back translation to be embedding into source language.
    \begin{itemize}
    \item Interpreter for back translation is not able to accurately replicate code length.
    \end{itemize}
  \end{itemize}
\item Worlds similar to CSF paper, but invariants on seals as well.
  \begin{itemize}
  \item Invariants says what can be sealed with a certain seals.
  \end{itemize}
\item Linker: resolves symbols
  \begin{itemize}
  \item Export refs (Question: do we allow other things than sealed code/data pairs to be exported?)
  \item Import refs
  \item Fresh seal requirement
  \end{itemize}
\item Components are memory segments and symbols.
\item Variable instruction length in the source to avoid leaking information through code size that cannot be matched after compilation (breaks compiler correctness, see example in subsection~\ref{subsec:capability-opacity}).
  \begin{itemize}
  \item Notice x86 allows instructions of up to infinite length, so this is not weird.
  \end{itemize}
\item Have opaque capabilities.
  \begin{itemize}
  \item Without opaque capabilities, stack consumption could be inferred through stack pointer index etc.
  \end{itemize}
\item Considered PORs for transition systems to: ($a \rightarrow b \Rightarrow a \rightarrow a \text{ and } b \rightarrow b$)
  \dominique{I don't remember what we meant with PORs, but perhaps quasi-reflexive partial orders.  Not sure what the point of this was again... }
  \begin{itemize}
  \item Define one region that governs the stack
  \item Make the regions more general
  \item For now, we have not done this: instead we stick with the transition systems we have and have two regions governing the stack reflecting the CSF work.
  \end{itemize}
\end{enumerate}

>>>>

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
\lau{13-09-2017: For now I have added linearity to the seals.}
\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
\lau{28-09-2017: Why would you like this? If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail. We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part. Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}. This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame. Makes sense?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

The source machine is also a capability machine with memory capabilities and sealed capabilities. Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    n & \in & \nats \\
    \src{r} &\in &  \RegName\\
    \src{\rn} &\defbnf & \src{r} \mid n \\
    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
  \end{array}
\]
\lau{30-08-2017: We had $\rret$ as not being one of the registers on the source machine. I could not recall why we had this, so I removed it. (I think it may have been from when we had call and return instructions?)}
\dominique{26-9-2017: I think we should allow $\rstk$ in the source language too.}
\lau{28-09-2017: I can see from the commit history that this was something when we had push/pop instructions as well as a return instruction. I have removed the restriction on $\rstk$.}
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.

@@

\dominique{27-09-2017: I think we need to do something about tail calls in the source language?}
\lau{27-09-2017: In what sense? Do you want to allow them (I do not thinkn they are allowed at the moment?)}

@@

\section{Compiler}
\[
\comp{\cdot} : \src{i}^* \fun \trg{i}^*
\]
\dominique{27-09-2017: Regarding the type signature for $[\cdot]$ above: does this mean that it maps lists of instructions to lists of instructions? Is there another function that compiles a single instruction to a list of instructions?}
\lau{28-09-2017: Yes, that was the thought at the moment, but this may very well change soon. This compiler would then use a compiler that compiles single instructions. That said, this part was added a long time ago and for some reason (some details about the compilation probably did not make sense to me at the moment), I did not complete it. I think the compiler will probably change to be from programs to programs, and a programs is not just going to be a list of instructions.}

@@

\dominique{27-09-2017: Is this example still relevant? Isn't it solved differently by the variable-length-instructions idea?}
\lau{28-09-2017: It seems to me that this is indeed solved by the variable length instructions. That said, I would like to keep it around (maybe change it so it becomes an argument for variable length instructions).}
This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{}{}$ and $\sreturn$ instructions. The below example is a motivation for having variable length instructions in the source language because if we have enough memory to ``do the same'' in the target language as in the source language, then the below example does not work.

@@

\section{Notes}
\subsection{Leuven stay conclusions}
\begin{enumerate}
\item Enter capabilities replaces by sealed code/data pairs \label{item:first-point}
  \begin{itemize}
  \item To allow us to forbid dynamic code generation
  \end{itemize}
\item Conditional full-abstraction
  \begin{itemize}
  \item No undefined behaviour by \emph{trusted code} in any context implies full-abstraction. (Blame idea: undefined as undef of current pc, so current pc can be checked for whether it was the trusted code).
  \item Avoids dynamic checks to protect trusted code against itself.
  \item Possible due to point \ref{item:first-point}.
  \end{itemize}
\item replace push/pull/call/ret/sload by symbolic return pointer (pair) and stack pointer.
  \begin{itemize}
  \item Allows back translation to be embedding into source language.
    \begin{itemize}
    \item Interpreter for back translation is not able to accurately replicate code length.
    \end{itemize}
  \end{itemize}
\item Worlds similar to CSF paper, but invariants on seals as well.
  \begin{itemize}
  \item Invariants says what can be sealed with a certain seals.
  \end{itemize}
\item Linker: resolves symbols
  \begin{itemize}
  \item Export refs (Question: do we allow other things than sealed code/data pairs to be exported?)
  \item Import refs
  \item Fresh seal requirement
  \end{itemize}
\item Components are memory segments and symbols.
\item Variable instruction length in the source to avoid leaking information through code size that cannot be matched after compilation (breaks compiler correctness, see example in subsection~\ref{subsec:capability-opacity}).
  \begin{itemize}
  \item Notice x86 allows instructions of up to infinite length, so this is not weird.
  \end{itemize}
\item Have opaque capabilities.
  \begin{itemize}
  \item Without opaque capabilities, stack consumption could be inferred through stack pointer index etc.
  \end{itemize}
\item Considered PORs for transition systems to: ($a \rightarrow b \Rightarrow a \rightarrow a \text{ and } b \rightarrow b$)
  \dominique{27-09-2017: I don't remember what we meant with PORs, but perhaps quasi-reflexive partial orders.  Not sure what the point of this was again... }
  \lau{28-09-2017: I think the next three points is the answer. We basically wanted a formalism in which we could design one region that would govern the stack. I do not remeber why it was made possible with PORs.}
  \begin{itemize}
  \item Define one region that governs the stack
  \item Make the regions more general
  \item For now, we have not done this: instead we stick with the transition systems we have and have two regions governing the stack reflecting the CSF work.
  \end{itemize}
\end{enumerate}
